---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r data}
dir.create("data")
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", 
              "./data/download.zip")
unzip("./data/download.zip", exdir = "./data")
activity <- read.csv("./data/activity.csv")
library(dplyr)
```

## What is mean total number of steps taken per day?

```{r mean}
activity <- activity %>% group_by(date)
asm <- activity %>% summarise(total_steps = sum(steps))
mean(asm$total_steps, na.rm = TRUE)
median(asm$total_steps, na.rm = TRUE)
```

## What is the average daily activity pattern?

```{r activity}

activity <- activity %>% group_by(interval)
activity.comp <- complete.cases(activity)
activity.c <- activity[activity.comp,]
asi <- activity.c %>% summarise(total_steps = sum(steps))
plot(asi$interval, asi$total_steps, type = "l", col = "red")
asi[asi$total_steps == max(asi),]

```

## Imputing missing values

```{r missing values}
sum(is.na(activity))
activity.m <- activity %>% group_by(interval)
adm <- activity.m %>% summarise(steps_mean = mean(steps, na.rm = TRUE))
adm$steps_mean <- c(round(adm$steps_mean))
 for (i in 1:nrow(activity.m)){
   if (isTRUE(is.na(activity.m[i,1]))) {
     activity.m[i,1] <- c(adm[adm$interval == activity.m$interval[i],2])
   }
 }

activity.m <- activity.m %>% group_by(date)
ams <- activity.m %>% summarise(steps_sum = sum(steps))
mean(ams$steps_sum)
median(ams$steps_sum)

hist(ams$steps_sum)


```

The impact of removing NA's from the orginal dataset was a decrease in both mean and median values when compared to the dataset where NA's were replaced with rounded mean values.


## Are there differences in activity patterns between weekdays and weekends?


```{r weekdays}
library(lubridate)
activity.m <- bind_cols(activity.m,
                        as.character(c(rep(NA,nrow(activity.m)))))
names(activity.m)[4] <- "day_type"
activity.m$date <- ymd(activity.m$date)
wdv <- unique(weekdays(activity.m$date))[1:5] 
for (i in 1:nrow(activity.m)){
   if (weekdays(activity.m[[i,"date"]]) %in% wdv) {
      activity.m[[i,"day_type"]] <- c("weekday")
   }else{activity.m[[i,"day_type"]] <- c("weekend")}
}
activity.m$day_type <- as.factor(activity.m$day_type)
```

```{r weekday timeseries}
library(lattice)
activity.m <- activity.m %>% group_by(interval,day_type)
ams <- activity.m %>% summarise(steps_mean = sum(steps))
xyplot(steps_mean ~ interval | day_type, data = ams, type = "l")

```
